---
title: "Preparing Files for Research Methods"
author: "DARS"
date: "6/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r, message = FALSE}
library(tidyverse)
library(readxl)
```
#Preparing Research Methodology file
##Import all courses from catalogues 
Source: DARS
```{r import old}
load("~/Documents/DARS/Fresh DARS/DARS/Output/d_text.RDATA")
```
##Import new courses Source: 
Edith- ILOs
```{r import new}
#import raw
new_courses_ilo <- read_xlsx("~/Documents/Summer UCM/ILO/ILO_2018_2019.xlsx")

#clean without ilo
new_courses     <- new_courses_ilo %>% select(Course) %>% distinct() 
new_courses
```
##Get all courses vector:
I get all the courses from the UCM catalogues, as this is the most comprehensive list available:
```{r}
courses <- d_text$catalogues %>% 
  filter(department == "UCM") %>% 
  select(`Course ID`) %>% 
  distinct()
courses
```
I check whether all new courses are included:
```{r which new courses are already in data}
courses_new_included <- tibble(Course = new_courses$Course) %>%
  mutate(Included = case_when(Course %in% courses$`Course ID`~T,
         T~F))
#Check to see how many are included and not included:
table(courses_new_included$Included)
```
Of the new courses only UGR2001 is not included in the previous course catalogues, this is probably just a name issue. 
```{r missing from old}
courses_new_included %>% filter(Included == F)
```
To be thorow, let us check which old courses are not in the new courses:
```{r}
courses_old_included <- tibble(Course = courses$`Course ID`) %>%
  mutate(Included = case_when(Course %in% new_courses$Course~T,
         T~F))
#Check to see how many are included and not included:
table(courses_old_included$Included)
```
We have 32 discontinued courses:
```{r discontinued}

```

```{r}
research <- rep(courses$`Course ID`, each = 14)
research_methodology <- tibble(`Course ID` = research) %>% 
  mutate(Task = rep(1:14, times = length(`Course ID`)/14))

write_excel_csv(courses,path = "~/Documents/Summer UCM/Methodology/ResearchMethodologies.csv")
```

