---
title: "Preparing Files for Research Methods"
author: "DARS"
date: "6/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r, message = FALSE}
library(tidyverse)
library(readxl)
```
#Preparing Research Methodology file
##Import all courses from catalogues 
Source: DARS
```{r import old}
load("~/Documents/DARS/Fresh DARS/DARS/Output/d_text.RDATA")
```
##Import new courses Source: 
Edith- ILOs
```{r import new}
#import raw
new_courses_ilo <- read_xlsx("~/Documents/Summer UCM/ILO/ILO_2018_2019.xlsx")

#clean without ilo
new_courses     <- new_courses_ilo %>% select(Course) %>% distinct() 
new_courses
```
##Get all courses vector:
I get all the courses from the UCM catalogues, as this is the most comprehensive list available:
```{r}
courses <- d_text$catalogues %>% 
  filter(department == "UCM") %>% 
  select(`Course ID`) %>% 
  distinct()
courses
```
I check whether all new courses are included:
```{r which new courses are already in data}
courses_new_included <- tibble(Course = new_courses$Course) %>%
  mutate(Included = case_when(Course %in% courses$`Course ID`~T,
         T~F))
#Check to see how many are included and not included:
table(courses_new_included$Included)
```
Of the new courses only UGR2001 is not included in the previous course catalogues, this is probably just a name issue. 
```{r missing from old}
courses_new_included %>% filter(Included == F)
```
To be thorow, let us check which old courses are not in the new courses (new courses are ILO courses for which we have the most uptodate data):
```{r}
courses_old_included <- tibble(Course = courses$`Course ID`) %>%
  mutate(Included = case_when(Course %in% new_courses$Course~T,
         T~F))
#Check to see how many are included and not included:
table(courses_old_included$Included)
```
We have 32 'discontinued' courses, but these include all projects because we were not given the file. 
```{r discontinued}
courses_old_included %>% filter(Included == F)
```
Okay, lets get ALL courses in a single data frame, and mark which are not in our ILO file, and which are not in our old files. We suspect some of the not_in_ILO file courses to be discontinued, we might not find course manuals for them:
```{r all courses}
courses_all <- tibble(Course = courses$`Course ID`) %>%
  full_join(new_courses, by = "Course") %>%
  mutate(included_ILO_file = ifelse(Course %in% new_courses$Course, T,F),
         included_in_Old_data = ifelse(Course %in% courses$`Course ID`, T,F))

courses_all
```
#Research Methodology Document (Tasks and articles)
Now, we would like to see what research methodology is used in each course, per task. For this we need a spreadhseet with all courses split into tasks. Generally there are 12 tasks but to be on the save side, lets say there are 14 tasks. 
We also want to be able to track the methodology to each article. I will assume that there are 10 articles per task.

This is going to be a huge spreadsheet:
```{r research spreadsheet}
course_codes <- courses_all$Course
tasks        <- 1:14
articles     <- 1:10

research_methodology <- expand.grid(articles, tasks, course_codes) %>% 
  rename(Article = Var1, Task = Var2, Course = Var3) %>%
  select(Course, Task, Article)


write_excel_csv(research_methodology, path = "~/Documents/Summer UCM/Methodology/ResearchMethodologies_Articles.csv")

research_methodology
```

We open the csv on excel and save it as an excel sheet

#Research Methodology Document (Taught Courses)
I'll create a second research methodology spreadsheet for every course
```{r}
rm_past_present <- tibble(Course = course_codes)

write_excel_csv(rm_past_present, path = "~/Documents/Summer UCM/Methodology/ResearchMethodologies_courses.csv" )

```

